Code walkthrough

### 2022-10-06 src/explore_aghq/walkthrough.Rmd
Start working on aghq::sample_marginal walkthrough

class(quad) <- c("marginallaplace", "aghq")
```

>>>>

class(quad) <- c("marginallaplace", "aghq")

```

# `aghq::sample_marginal`

What we'd like to do is to sample from
$$
\begin{align}
\tilde p(x_i \, | \, y) &= \sum_z \tilde p_\text{G}(x_i \, | \, \theta(z), y) \tilde p_\text{LA}(\theta(z) \, | \, y) w_k(z), \\
&= \sum_z \tilde p_\text{G}(x_i \, | \, \theta(z), y) \lambda(z)
\end{align}
$$
where $\lambda(z) \in [0, 1]$ are weights such that $\sum_z \lambda(z) = 1$.
We can do this by

1. Sample $z_j \sim \text{Multinomial}(1, \lambda(z_1), \ldots, \lambda(z_J))$
2. Given $z_j$ sample from $\tilde p_\text{G}(x \, | \, \theta(z_j), y)$
3. Take the $i$th entry of $\tilde p_\text{G}(x \, | \, \theta(z_j), y)$ as the sample from $\tilde p_\text{G}(x_i \, | \, \theta(z_j), y)$

Now that we have `quad`, it should be compatible with `sample_marginal`:

```{r}
#' The number of samples to get
M <- 10

transformation <- quad$transformation
interpolation <- "auto"

#' Is this getting mc.cores, and if there isn't a value just setting it to 1 (integer)
numcores <- getOption("mc.cores", 1L)

#' Length of the random effects
d <- dim(quad$modesandhessians$H[[1]])[1]

simlist <- quad$modesandhessians

#' Avoid use of parallel computing on Windows
if (.Platform$OS.type == 'windows') numcores <- 1
```

The next step is to compute the Cholesky decomposition `L` of the Hessian `H` such that
$$
H = LL^\top
$$
The reason we want the Cholesky is that we will be using it to same from the joint Gasusian approximation.
The function `chol` from `base` will compute it for us.
For example, for the first element:

```{r}
h <- simlist$H[[1]]
h <- Matrix::forceSymmetric(h)
L <- chol(h, perm = FALSE)
L
```

Doing this for the list using `lapply`:

```{r}
simlist$L <- lapply(simlist$H,function(h) chol(Matrix::forceSymmetric(h), perm = FALSE))
```

There is another version, using the arguments `perm` and `LDL` (which I don't understand) that may be quicker but ultimately will achieve the same thing:

```{r}
# simlist$L <- lapply(simlist$H, function(h) Matrix::Cholesky(as(Matrix::forceSymmetric(h), "sparseMatrix"), perm = TRUE, LDL = FALSE))
```

Now let's compute the $\lambda(z)$ weights.

```{r}
simlist$lambda <- exp(quad$normalized_posterior$nodesandweights$logpost_normalized) * quad$normalized_posterior$nodesandweights$weights
simlist$lambda
```

We can check that they do indeed sum to one:

```{r}
sum(simlist$lambda)
```

Now let's sample from the multinomial:

```{r eval=FALSE}
j <- apply(stats::rmultinom(M, 1, simlist$lambda), 2, function(x) which(x == 1))
j

#' Values are number of samples to draw for each j
#' In this case they're likely to be all 1, since we're not doing many samples
tt <- table(j)
```

Now we construct a big Gaussian mixture matrix.
The first step is to create a matrix with `d` columns, one for each random effect, and `M` rows, one for each sample, where each value is a standard normal $\mathcal{N}(0, 1)$:

```{r}
A <- matrix(stats::rnorm(M * d), nrow = M)

#' Now make a list where each element is a row of A, where the index is determined by j
A_split <- split(A, j)

#' Big Gaussian mixture matrix

#' For each element of the list, create a matrix with d rows (and 1 column? always?)
Z <- lapply(A_split, matrix , nrow = d)
```
